Postural changes mediate children's visual access to social information

### 2018-01-13 analysis-main/main_analyses_bll.Rmd
discussion intro integrated

```{r}
d <- read_csv("../data/consolidated_data_3dets.csv") %>%
  mutate(posture = factor(posture), 
         orientation = factor(orientation))
demo <- read_csv("../data/demographics/demographics.csv")
#cdis <- read_csv("../data/cdis/cdi_summary.csv") ## not in repo
```

@@

```{r}
# mss <- d %>%
#   filter(!is.na(posture)) %>%
#   filter(!is.na(face)) %>%
#   group_by(age.grp, posture, subid) %>%
#   summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
#   group_by(age.grp)
# 
# 
# sb_summary <- sb %>%
#   filter(!is.na(posture)) %>%
#   filter(!is.na(face)) %>%
#   group_by(age.grp, posture, subid) %>%
#   dplyr::summarise(face = count(face)) %>%
#   group_by(age.grp)
# 
# 
# nameFreq <- sb %>%
#   filter(!is.na(posture)) %>%
#   filter(!is.na(face)) %>%
#   group_by(age.grp, posture, subid) %>%
#   dplyr::summarise(nameFreq= max(naming.instance, na.rm="TRUE")) %>%
#   group_by(age.grp)
# 
# ggplot(nameFreq, aes(x = factor(age.grp), y = nameFreq, col = posture)) + 
#   geom_boxplot() + 
#   ylab("within window") + 
#   xlab("Age (months)") + 
#   theme_bw() 
# 
# 
# # ggplot(sb_summary, aes(x = factor(age.grp), y = face, col = posture)) + 
# #   geom_boxplot() + 
# #   ylab("within window") + 
# #   xlab("Age (months)") + 
# #   theme_bw()

sb <- sb %>%
  filter(!is.na(posture))

ggplot(sb, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot() + 
  ylab("within window") + 
  xlab("Age (months)") + 
  facet_grid(.~familiarity) +
  theme_bw()

@@

```{r}
  qplot(age.grp,face,colour=window,
        xlab="Age (Months)",ylab="Proportion faces detected in window",
        data=mss,geom=c("point")) +
    scale_x_continuous(breaks=c(8,12,16)) +
    geom_smooth()
```

>>>>

```{r}
d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
  mutate(posture = factor(posture), 
         orientation = factor(orientation))
demo <- read_csv("../data/demographics/demographics.csv")
#cdis <- read_csv("../data/cdis/cdi_summary.csv") ## not in repo
```

@@

@@

<!-- # mss <- d %>% -->
<!-- #   filter(!is.na(posture)) %>% -->
<!-- #   filter(!is.na(face)) %>% -->
<!-- #   group_by(age.grp, posture, subid) %>% -->
<!-- #   summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>% -->
<!-- #   group_by(age.grp) -->
<!-- #  -->
<!-- #  -->
<!-- # sb_summary <- sb %>% -->
<!-- #   filter(!is.na(posture)) %>% -->
<!-- #   filter(!is.na(face)) %>% -->
<!-- #   group_by(age.grp, posture, subid) %>% -->
<!-- #   dplyr::summarise(face = count(face)) %>% -->
<!-- #   group_by(age.grp) -->
<!-- #  -->
<!-- #  -->
<!-- # nameFreq <- sb %>% -->
<!-- #   filter(!is.na(posture)) %>% -->
<!-- #   filter(!is.na(face)) %>% -->
<!-- #   group_by(age.grp, posture, subid) %>% -->
<!-- #   dplyr::summarise(nameFreq= max(naming.instance, na.rm="TRUE")) %>% -->
<!-- #   group_by(age.grp) -->
<!-- #  -->
<!-- # ggplot(nameFreq, aes(x = factor(age.grp), y = nameFreq, col = posture)) +  -->
<!-- #   geom_boxplot() +  -->
<!-- #   ylab("within window") +  -->
<!-- #   xlab("Age (months)") +  -->
<!-- #   theme_bw()  -->
<!-- #  -->
<!-- #  -->
<!-- # # ggplot(sb_summary, aes(x = factor(age.grp), y = face, col = posture)) +  -->
<!-- # #   geom_boxplot() +  -->
<!-- # #   ylab("within window") +  -->
<!-- # #   xlab("Age (months)") +  -->
<!-- # #   theme_bw()  -->

<!-- sb <- sb %>% -->
<!--   filter(!is.na(posture))  -->

<!-- ggplot(sb, aes(x = factor(age.grp), y = face, col = posture)) +  -->
<!--   geom_boxplot() +  -->
<!--   ylab("within window") +  -->
<!--   xlab("Age (months)") +  -->
<!--   facet_grid(.~familiarity) + -->
<!--   theme_bw()  -->

<!-- ``` -->

<!-- ```{r} -->
<!--   qplot(age.grp,face,colour=window, -->
<!--         xlab="Age (Months)",ylab="Proportion faces detected in window", -->
<!--         data=mss,geom=c("point")) + -->
<!--     scale_x_continuous(breaks=c(8,12,16)) + -->
<!--     geom_smooth() -->
<!-- ``` -->

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
discussion intro integrated

---
title: "Postural changes mediate children's visual access to faces and hands"
bibliography: xsface.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

abstract:

@@

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(plyr) # need to load this for scripts, always load before
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(lme4)
```

@@

\begin{table}[t]
  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
  \begin{center}
    \begin{tabular}{l|ccc|ccc}
      \hline
       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
      % \null Model & P & R & F & P & R & F  \\
      \hline
      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose & .86 & .78 & .81 & .93 & .76 & .83 \\
    \hline
    \end{tabular}
  \end{center}
\end{table}

@@

```{r}
d <- read_csv("../data/consolidated_data.csv") %>%
  mutate(posture = factor(posture), 
         orientation = factor(orientation))

facebyAge <- d %>%
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test) %>%
  summarise(face = mean(face)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len_seconds/60))

faceByOrientation <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

facebyPosture <- d %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, posture, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)
```

@@

```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
ggplot(facebyAge, aes(x = age.at.test, y = face)) + 
  geom_jitter(width = .25, aes(size = len)) + 
  geom_smooth() + 
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_size_continuous(name = "Length (s)") + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw()
```

```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
ggplot(faceByOrientation, aes(x = factor(age.grp), y = face, col = orientation)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Model
```{r}

fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), 
      data = d,  
      family = "binomial")

kable(summary(mod_covariates_age_v2)$coef, digits = 3)

@@

@@

>>>>

---
title: "Postural changes mediate children's visual access to social information"
bibliography: xsface.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

@@

abstract: 
The aiblity to rpocess social cues (e.g., eye gaze) is a critical component in children/s early language development. However, As children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way. How do these postural changes affect children's access to the social information relevant for word-learning?

@@

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(lme4)
```
# Introduction
Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life. However, children do not learn words in a vacuum but rather in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process. Further, children’s ability to effectively process these social cues may be a key factor in their early language development.  In a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).  More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months (@brooks2005) and throughout the second year of life (@brooks2008).

However, as children are learning their first words, their view of the world is also changing radically [@adolph2007]. Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see. During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).

One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities (Iverson, 2010). Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”).  (Karasik et al., 2014).  More directly, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.  On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting. In other words, the ability to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.

Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; @franchak2011, @yoshida2008; @clerkin2017). Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017]. Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2017].

Here, we directly ask whether the postural changes that infants’ experience as they reach their first birthday change the availability of the social information relevant for word learning. To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated. We capitalize on recent improvements in computer vision algorithms (@citemtcnn, @openpose) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.

@@

\begin{table}[t]
  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
  \begin{center}
    \begin{tabular}{l|ccc|ccc}
      \hline
       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
      % \null Model & P & R & F & P & R & F  \\
      \hline
      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose-Faces & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose-Wrists & .86 & .78 & .81 & .93 & .76 & .83 \\
      % OpenPose-Hands & .86 & .78 & .81 & .93 & .76 & .83 \\
    \hline
    \end{tabular}
  \end{center}
\end{table}

@@

```{r}
d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = face, faceMT, faceOP)

facebyAge <- dlong %>%
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test, detectorType) %>%
  summarise(face = mean(face)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len_seconds/60))

faceByOrientation <- dlong %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid, detectorType) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

facebyPosture <- dlong %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, posture, subid,detectorType) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)
```

@@

```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }

  ggplot(facebyAge, aes(x = age.at.test, y = face, color = detectorType, size = len)) +
    geom_jitter(width = .25) + 
    geom_smooth() +
    scale_x_continuous(breaks=c(8,12,16)) +
    scale_size_continuous(range=c(1,4),name = "Video Length (mins)") + 
    ylab("Proportion Face Detections") + 
    xlab("Age (months)") + 
    theme_bw() +
    theme(legend.position="bottom", legend.box = "horizontal")

@@

```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)"}
ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  facet_grid(~detectorType) +
  theme_bw()

```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)" }
ggplot(faceByOrientation, aes(x = factor(age.grp), y = face, col = orientation)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  facet_grid(~detectorType) +
  theme_bw() 
  # theme(legend.position="bottom", legend.box = "horizontal")

<!-- ## Model -->

<!-- # ```{r} -->
<!-- # d$faceNum=as.numeric(d$faceMT) -->
<!-- # -->
<!-- # # Fails to converge -->
<!-- # # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
<!-- # #       data = d, -->
<!-- # #       family = "binomial") -->
<!-- # -->
<!-- # # Fails to converge -->
<!-- # # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
<!-- # #       data = d, -->
<!-- # #       family = "binomial") -->
<!-- # -->
<!-- # fullModelminsInt <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
<!-- #        data = d, -->
<!-- #        family = "binomial") -->
<!-- # -->
<!-- # ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid), -->
<!-- #       data = d, -->
<!-- #       family = "binomial") -->
<!-- # -->
<!-- # ``` -->

To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame (as detected by the MTCNN3 model) as the dependent variable, and participant's age, orientation, and posture as predictors. Interactions between predictors were not included as this maximal model failed to converge. A summary of the coefficients of this model can be found in Table {ref}. While age remained a significant predictor even when accounting for the effects of infants' posture and caregiver's oirentation, it held relativley less predictive power. Overall, these results confirm that infant's access to faces is heavily infleunced by their own posture and their caregivers orientation towards them.

<!-- ```{r xtable, results="asis"} -->
<!-- tab2 <- xtable::xtable(summary(fullModelminsInt)$coef, digits=3,  -->
<!--                       caption = "Results from GLMM model prediting the presence/absence of a face across                       the entire dataset.") -->

<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
<!-- ``` -->

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model prediting the presence/absence of a face across                       the entire dataset.} 
\end{table}

@@

Next, we examined whether we would see similar trends in infant and children's access to hands.

# General Discussion
We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information relevant for word-learning, here operationalized as the presence of the faces and hands of their caregiver. We found that children's posture and orientation towards their caregiver changed systematically across age, and that all three of these factors dramatically impacted the proportion of faces and hands that were avaliable in the child's visual field. Thus, infants postural development is mediating factor that explains age-related changes in the proportion of faces and hands avaliable in infants' visual field.

This work also deploys novel advancements in computer vision to the study of developmental psychology. The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms (Krihevsky et al., 2012), creating a new generation of algorithmic tools that both are substantially better equipped to deal with noisier, more complicated datasets and that can  extract richer and more detailed information. Indeed, these videos from the infant perspective provided sustantial challenges (e.g., partially occluded faces) for the classc models of face detedction (e.g., viola jones).   Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to new infant learners.

Thus, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspecive on the visual world and the implications of these changes for both linguistic, cognitive, and social development.
